### 2024-02-20 R/processing-code/Processing-Quarto-Output.qmd

---
title: "Data Processing"
author: "Kelly Cao and Rachel Robertson"
date: "2023-01-03"
output: html_document
editor: 
  markdown: 
    wrap: sentence
---

# Setup

This chunk pulls the R-script entitled Creating_Data_Frame.R.

```{r, include=FALSE, cache=FALSE}
knitr::read_chunk('processing-code/creating-df/Creating_Data_Frame.R')
```

```{r, include=FALSE, cache=FALSE}
knitr::read_chunk('R/processing-code/Cleaning Data/cleaning1.R')
```

Load needed packages.

```{r, packages, message = FALSE, warning = FALSE}
```

# Data loading

Note that for functions that come from specific packages (instead of base R), I often specify both package and function like so: package::function() that's not required one could just call the function specifying the package makes it clearer where the function "lives", but it adds typing.
You can do it either way.

```{r, loaddata}
```

# Specifying Factor Variables

By inspecting the data as done above, we find some problems that need addressing:

First, there is an entry for height which says "sixty" instead of a number.
Does that mean it should be a numeric 60?
It somehow doesn't make sense since the weight is 60kg, which can't happen for a 60cm person (a baby).
Since we don't know how to fix this, we might decide to remove the person.
This "sixty" entry also turned all Height entries into characters instead of numeric.
That conversion to character also means that our summary function isn't very meaningful.
So let's fix that first.

```{r, cleandata1}
```

Now we see that there is one person with a height of 6.
That could be a typo, or someone mistakenly entered their height in feet.
Since we unfortunately don't know, we might need to remove this person, which we'll do here.

# Save data

All done, data is clean now.
Time to save.
Let's assign at the end to some final variable, this makes it easier to add further cleaning steps above.

Finally, we save the clean data as RDS file.
I suggest you save your processed and cleaned data as RDS or RDA/Rdata files.
This preserves coding like factors, characters, numeric, etc.
If you save as CSV, that information would get lost.
However, CSV is better for sharing with others since it's plain text.
If you do CSV, you might want to write down somewhere what each variable is.

See here for some suggestions on how to store your processed data: http://www.sthda.com/english/wiki/saving-data-into-r-data-format-rds-and-rdata

```{r, savedata}
```

# Cleaning Data

```{r, loaddata2}
```

## Exploring the data

```{r, exploredata2}
```

## Eliminating NA Values and Columns

```{r, cleandata2}
```

## Saving the Data Frame

```{r, savedata2}
```