### 2024-02-21 R/eda-code/EDA-code.qmd
Created edacode.R and eda-code.qmd. Uploaded some tables and figures in /result directory

---
title: "Exploratory Analysis on CDC Teen Vaccination Surveys"
date: "`r format(Sys.Date()'"
Author: Kelly Cao and Rachel Robertson
output: html_document
editor: 
  markdown: 
    wrap: sentence
---

The following file will load the cleaned data and perform some exploration.

The categorical and numerical variables of interest will be plotted and placed in models for analysis.

## Loading the .R script which performed the analysis

```{r, include=FALSE, cache=FALSE}
library(here)
knitr::read_chunk(here("./R/eda-code/edacode.R"))
```

## Loading the Packages

```{r, packages, message = FALSE, warning = FALSE}
```

```{r, loaddata}
```

## Showing Summary of Overall Data Set

The following data set has been cleaned. Any observations with NAs has been removed. 
```{r, table1}

```

## Addressing Numerical Data Set
The distribution of the numerical income to poverty ratio can be shown below. It displays a value from 1-3. A ratio displaying less than 1 depicts an income less than the poverty level. Any income ratio >1 indicates an income greater than the poverty level. Most of the observations were gathered from households with income ratios of 3, as shown by the skewed plot below.

```{r, INCPORAR_I}

```

## Addressing Categorical Data

The state variable shows a more uniformed distribution, suggesting that the observations were gathered fairly across the US states and territory. Although it is noted a few territories provides 0 observations.

```{r, STATE}

```

The family income variable provides a more normal distribution with the majority of the observation coming from $75000 income level.

```{r, INCQ298A}

```

The race and ethnicity variables show a normal distribution.
```{r, RACEETHK}

```

```{r, RACE_K}

```

The Insurance status variable shows that the majority of the observations came from households with private insurance only. This is mirrored in the facility variable, where a majority of its observation s show private facilities being the location.

```{r, INS_STAT2_I}

```

```{r, FACILITY}

```

The vaccination variable indicates nearly 80% of the observations had an Up-to-date status for the 1+ shot HPV vaccination.

```{r, Vaccination}

```

## Model and Plot Fitting

### Fit 1

A boxplot is created with income-poverty ratio to vaccination status. The box plot indicates little difference between the two status. A statistical model is fitted to the same variables. A p-value of 0.1358 suggests there is little significance between the income-poverty variable and the vaccination status.

```{r, fitfig1}

```

### Fit 2
A boxplot is created with income-poverty ratio to insurance status. The box plot indicates there are distinct difference between the two status. A statistical model is fitted to the same variables. The p-value of < 2.26e-16 suggests there is statistical significance between the income-poverty variable and the insurance status.

```{r, fitfig2}

```