### 2024-02-22 code/eda-code/EDA-code.qmd
Added to EDA, staged in Manuscript
I added 5 figures/ tables to the EDA and in addition I added the EDA chunk to the manuscript document. I would prefer to just add the figure images preceding the description, but I cannot get markdown to allow me to pull just the figures.

---
title: "Exploratory Analysis on CDC Teen Vaccination Surveys"
date: "`r format(Sys.Date()'"
Author: Kelly Cao and Rachel Robertson
output: html_document
editor: 
  markdown: 
    wrap: sentence
---

The following file will load the cleaned data and perform some exploration.

The categorical and numerical variables of interest will be plotted and placed in models for analysis.

## Loading the .R script which performed the analysis

```{r, include=FALSE, cache=FALSE}
library(here)
knitr::read_chunk(here("./code/eda-code/edacode.R"))
```

## Loading the Packages

```{r, packages, message = FALSE, warning = FALSE}
```

```{r, loaddata}
```

## Showing Summary of Overall Data Set

The following data set has been cleaned.
Any observations with NAs has been removed.

```{r, table1}

```

## Addressing Numerical Data Set

The distribution of the numerical income to poverty ratio can be shown below.
It displays a value from 1-3.
A ratio displaying less than 1 depicts an income less than the poverty level.
Any income ratio \>1 indicates an income greater than the poverty level.
Most of the observations were gathered from households with income ratios of 3, as shown by the skewed plot below.

```{r, INCPORAR_I}

```

## Addressing Categorical Data

The state variable shows a more uniformed distribution, suggesting that the observations were gathered fairly across the US states and territory.
Although it is noted a few territories provides 0 observations.

```{r, STATE}

```

The family income variable provides a more normal distribution with the majority of the observation coming from \$75000+ income level.

```{r, INCQ298A}

```

The race and ethnicity variables show a distribution that is similar to the U.S. census data, which concludes that the racial composition of the U.S. is 58.9% white, 13.6% African American, and 19.1% Hispanic or Latino.

```{r, RACEETHK}

```

```{r, RACE_K}

```

The Insurance status variable shows that the majority of the observations came from households with private insurance only. This is followed by any medicaid and then other insurance.
This is mirrored in the facility variable, where a majority of its observations show private facilities being the location.

```{r, INS_STAT2_I}

```

```{r, FACILITY}

```

The vaccination variable indicates nearly 80% of the observations had an Up-to-date status for the 1+ shot HPV vaccination.

```{r, Vaccination}

```

We examine the percentage of up-to-date vaccination status by state.
```{r, state-vaccine}

```

```{r, facility-vaccine}

```

```{r, race-ethnicity-vaccine}

```
## Model and Plot Fitting

### Fit 1

A boxplot is created with income-poverty ratio to vaccination status.
The box plot indicates little difference between the two status.
A statistical model is fitted to the same variables.
A p-value of 0.1358 suggests there is little significance between the income-poverty variable and the vaccination status.

```{r, fitfig1}

```

### Fit 2

A boxplot is created with income-poverty ratio to insurance status.
The box plot indicates there are distinct difference between the two status.
A statistical model is fitted to the same variables.
The p-value of \< 2.26e-16 suggests there is statistical significance between the income-poverty variable and the insurance status.

```{r, fitfig2}

```

### Fit 3
```{r, fitfig3}

```

### Fit 4
```{r, fitfig4}

```