### 2024-02-22 code/processing-code/Processing-Quarto-Output.qmd

---
title: "Data Processing"
author: "Kelly Cao and Rachel Robertson"
date: "2023-01-03"
output: html_document
editor: 
  markdown: 
    wrap: sentence
---

# Setup

This chunk pulls the R-script entitled Creating_Data_Frame.R.

```{r, include=FALSE, cache=FALSE}
library(here)
knitr::read_chunk(here("./code/processing-code/creating-df/Creating_Data_Frame.R"))
```

Load needed packages.

```{r, packages, message = FALSE, warning = FALSE}
```

# Data loading

```{r, loaddata}
```

# Specifying Factor Variables

```{r, cleandata}

```

# Save data

```{r, savedata}
```

# Cleaning Data

```{r, include=FALSE, cache=FALSE}
library(here)
knitr::read_chunk(here("./code/processing-code/cleaningdata/cleaning1.R"))
```

```{r, packages2}
```

```{r, loaddata2}
```

## Exploring the data

```{r, exploredata2}
```

## Eliminating NA Values and Columns

```{r, cleandata2}
```

## Saving the Data Frame

```{r, savedata2}
```