### 2024-03-15 code/analysis-code/analysis1.qmd
Added to manuscript and analysis code document. There is a start in analysis1 qmd.
My computer is freaking out so I am going to try to perform some more diagnostics on the model I got with stepwise comparisons, in the morning.

---
title: "Data Analysis and Modelling of Teen Vaccination Surveys"
date: "`r format(Sys.Date()'"
Author: Kelly Cao and Rachel Robertson
output: html_document
editor: 
  markdown: 
    wrap: sentence
---
First, I will start by setting an equation for the generalized logistic regression model with all of the predictor variables intended for use.
```{r}
levels(mydata$P_UTDHPV)
mydata2 <- droplevels(mydata[!mydata$P_UTDHPV == 'Missing Data', ]) # I will do some brief cleaning and remove all rows with missing values for our response variable, 'P_UTDHPV', to perform a binomial regression
str(mydata2$P_UTDHPV) # Check for removal of missing values
fitall <- glm(P_UTDHPV ~ AGE + SEX + STATE + INS_STAT2_I + INCQ298A + INS_BREAK_I + INCPOV1 + RACEETHK + EDUC1 + LANGUAGE + MOBIL_1 + RENT_OWN + FACILITY, data = mydata2, family = binomial) # define the equation for the binomial logistic regression
```

Now, I will set the start of the step-wise comparison, using a model with an intercept only (or blank model).
```{r}
fitstart <- glm(P_UTDHPV ~ 1, data=mydata2, family = binomial)
```

Now I will start the step-wise comparison in both the forwards and backwards direction to allow the model to add or drop predictors depending on the responding AIC value.
```{r}
step(fitstart, direction = "both", scope = list(lower = ~1, upper = ~ AGE + SEX + STATE + INS_STAT2_I + INCQ298A + INS_BREAK_I + INCPOV1 + RACEETHK + EDUC1 + LANGUAGE + MOBIL_1 + RENT_OWN + FACILITY), data = mydata, family = binomial) # define the lower and upper limits of the step wise comparison 
```
The model chosen includes the predictors STATE, FACILITY, RACEETHK, AGE, EDUC1, INS_BREAK_I, INCPOV1, SEX, LANGUAGE, MOBIL_1, and INS_STAT2_I. 
Because of the extremely high residual value, I will perform some diagnostics to determine what is the issue. First, I will look at a residual plot to check for non linearity. Next, I will use a VIF analysis to check for multicollinearity. I am suspicious of co linearity as many of these socioeconomic predicts might be inherently related.

I consulted ChatGPT to ask how to perform a VIF analysis in R and was recommended the packages 'car' and 'carData' using the vif() function.
I will start by opening the libraries necessary.
```{r}
library(ggplot2)
library(car) # to perform VIF analysis
library(carData) # to look at
```

Next I will produce a residual plot
```{r}
model <- glm(glm(P_UTDHPV ~ AGE + SEX + STATE + INS_STAT2_I + INS_BREAK_I + INCPOV1 + RACEETHK + EDUC1 + LANGUAGE + MOBIL_1 + FACILITY, data = mydata2, family = binomial)
# Residual plot
residualPlot <- function(model) {
  ggplot(data = mydata2(fit = predict(model), residuals = resid(model)),
         aes(x = fit, y = residuals)) +
    geom_point() +
    geom_hline(yintercept = 0, linetype = "dashed", color = "red") +
    xlab("Fitted values") +
    ylab("Residuals") +
    ggtitle("Residual Plot")
}

# VIF diagnostics
vifDiagnostics <- function(model) {
  vif_values <- car::vif(model)
  print(vif_values)
}

# Create residual plot
residual_plot <- residualPlot(fitall)
print(residual_plot)

# Calculate VIFs
vif_diagnostics <- vifDiagnostics(model)
print(vif_diagnostics)

```
```