### 2024-03-16 code/analysis-code/analysis1.qmd
Added analysis2.qmd, added more to analysis1

---
title: "Data Analysis and Modelling of Teen Vaccination Surveys"
date: '`r format(Sys.Date())`'
Author: Kelly Cao and Rachel Robertson
output: html_document
editor: 
  markdown: 
    wrap: sentence
---
First, I will start by setting an equation for the generalized logistic regression model with all of the predictor variables intended for use.

The model aims to predict the vaccination status (P_UTDHPV) based on various predictor variables like age, sex, state, insurance status, income, race/ethnicity, education, language, etc. The response variable P_UTDHPV represents whether a teenager is up-to-date on HPV vaccination.
```{r}
#Path to data. Note the use of the here() package and not absolute paths
data_location <- here::here("data","processed-data","cleandata1.rds")

#load data
mydata <- readRDS(data_location)

levels(mydata$P_UTDHPV)
mydata2 <- droplevels(mydata[!mydata$P_UTDHPV == 'Missing Data', ]) # I will do some brief cleaning and remove all rows with missing values for our response variable, 'P_UTDHPV', to perform a binomial regression
str(mydata2$P_UTDHPV) # Check for removal of missing values
fitall <- glm(P_UTDHPV ~ AGE + SEX + STATE + INS_STAT2_I + INCQ298A + INS_BREAK_I + INCPOV1 + RACEETHK + EDUC1 + LANGUAGE + MOBIL_1 + RENT_OWN + FACILITY, data = mydata2, family = binomial) # define the equation for the binomial logistic regression
```

> A Stepwise model comparison is performed to iteratively add or remove predictor variables based on their significance in improving the model fit.

Now, I will set the start of the step-wise comparison, using a model with an intercept only (or blank model). 
```{r}
fitstart <- glm(P_UTDHPV ~ 1, data=mydata2, family = binomial)
```

Now I will start the step-wise comparison in both the forwards and backwards direction to allow the model to add or drop predictors depending on the responding AIC value. This is performed with the `step()` function, setting up the scope for the variables to consider.
```{r}
step(fitstart, direction = "both", scope = list(lower = ~1, upper = ~ AGE + SEX + STATE + INS_STAT2_I + INCQ298A + INS_BREAK_I + INCPOV1 + RACEETHK + EDUC1 + LANGUAGE + MOBIL_1 + RENT_OWN + FACILITY), data = mydata, family = binomial) # define the lower and upper limits of the step wise comparison 
```
The model chosen includes the predictors STATE, FACILITY, RACEETHK, AGE, EDUC1, INS_BREAK_I, INCPOV1, SEX, LANGUAGE, MOBIL_1, and INS_STAT2_I. 
Because of the extremely high residual value, I will perform some diagnostics to determine what is the issue. First, I will look at a residual plot to check for non linearity. Next, I will use a VIF analysis to check for multicollinearity. I am suspicious of co linearity as many of these socioeconomic predicts might be inherently related.

I consulted ChatGPT to ask how to perform a VIF analysis in R and was recommended the packages 'car' and 'carData' using the vif() function.
I will start by opening the libraries necessary.
```{r}
library(ggplot2)
library(car) # to perform VIF analysis
library(carData)
```

Next I will produce a residual plot. I asked ChatGPT how to use ggplot to produce a residual plot for a logistic regression function and was given the base code, which I modified to fit my analysis.
```{r}
# First, I will define the model given to me by the step wise comparison
model <- glm(P_UTDHPV ~ AGE + SEX + STATE + INS_STAT2_I + INS_BREAK_I + INCPOV1 + RACEETHK + EDUC1 + LANGUAGE + MOBIL_1 + FACILITY, data = mydata2, family = binomial)

# Define function for residual plot of logistic regression
residualPlotLogit <- function(model) {
  fitted_prob <- predict(model, type = "response") # Predicted probabilities
  residuals <- resid(model, type = "response") # Calculate residuals
  plot_data <- data.frame(fit = fitted_prob, residuals = residuals) # Create data frame

  # Plot
  ggplot(data = plot_data,
         aes(x = fit, y = residuals)) +
    geom_point() +
    geom_hline(yintercept = 0, linetype = "dashed", color = "red") +
    xlab("Fitted probabilities") +
    ylab("Residuals") +
    ggtitle("Residual Plot of Proposed Stepwise Model")
}

# Create the residual plot
residual_plot_logit <- residualPlotLogit(fitall)
print(residual_plot_logit)

```
The residual plot shows clear collinearity that is not represented by the current model. To find the predictors with collinnearity we will use a correlation matrix. This is because the VIF analysis will not run when multicollinearity is too high.
```{r}
factor_data <- model.matrix(~ STATE + FACILITY + RACEETHK + AGE + EDUC1 + INS_BREAK_I + INCPOV1 + SEX + LANGUAGE + MOBIL_1 + INS_STAT2_I, data = mydata2)

# Compute correlation matrix
correlation_matrix <- cor(factor_data)

# Find indices of correlation matrix where values are 1 or 0
indices <- which(!is.na(correlation_matrix) & correlation_matrix == 1 | correlation_matrix == 0, arr.ind = TRUE)

# Extract variable names corresponding to these indices
variable_pairs <- rownames(correlation_matrix)[indices[,1]]
variable_pairs <- cbind(variable_pairs, colnames(correlation_matrix)[indices[,2]])

# Print variable pairs with correlation values of 1 or 0
print(variable_pairs)

```
This shows that many levels from each of the variables show multicollineaity. This might be addressed by using a different model for analysis.