### 2024-04-04 code/analysis-code/analysis3.qmd
Demographics table and lasso
Demographics table was added to the manuscript and the first round of lasso was fitted. However, for some reason the estimate for all of the predictors chosen by LASSO were equal to zero. I am not sure if this is because the penalty is too high or why else. I will try and research this.

---
title: "Data Analysis and Modelling of Teen Vaccination Surveys"
date: '`r format(Sys.Date())`'
Author: Kelly Cao and Rachel Robertson
output: html_document
editor: 
  markdown: 
    wrap: sentence
---
# LASSO using Tidymodels
We will start by openning the libraries that we will need for creating a LASSO regression model for the data
```{r}
library(tidymodels) # use tidymodels framework
library(ggplot2) # producing visual displays of data
library(dplyr) # manipulating and cleaning data
library(here) # making relative pathways
library(glmnet) # for LASSO regression engine
library(doParallel) # for parallel processing
library(rsample) # for cross validation
```

Now, we will load the preprocessed data.
```{r}
# Load and preprocess data
data_location <- here::here("data","processed-data","cleandata1.rds")
mydata <- readRDS(data_location)

# Remove rows with missing values
mydata$SEX <- droplevels(mydata$SEX, exclude = c("DON'T KNOW", "MISSING IN ERROR", "REFUSED"))
mydata$INS_STAT2_I <- droplevels(mydata$INS_STAT2_I, exclude = "MISSING Data")
mydata$STATE <- droplevels(mydata$STATE, exclude = "Missing Data")
mydata$MOBIL_1 <- droplevels(mydata$MOBIL_1, exclude = c("DON'T KNOW", "MISSING IN ERROR", "REFUSED"))
mydata$FACILTY <- droplevels(mydata$FACILITY, exclude = "Missing Data")
mydata$P_UTDHPV <- droplevels(mydata$P_UTDHPV, exclude = "Missing Data")
```
Next, we will split the data into training and testing groups so that the LASSO regression has a comparison group.
```{r}
# Split data into training and testing datasets
set.seed(123) # seed for reproducibility
split_data <- initial_split(mydata, prop = 0.8) # 80% split for training/testing data
train_data <- training(split_data)
test_data <- testing(split_data)
```
Next, we will specify the LASSO recipe, or the regression formula. This will contain the predictors of interest used for previous analysis.
```{r}
# Create recipe containing predictors of interest
lasso_rec <- recipe( P_UTDHPV ~ AGE + SEX + STATE + INS_STAT2_I + INCQ298A + INS_BREAK_I + INCPOV1 + RACEETHK + EDUC1 + LANGUAGE + MOBIL_1 + RENT_OWN + FACILITY, data = train_data) %>%
  step_dummy(all_nominal(), -all_outcomes())  %>%
  step_rm(all_outcomes(), -all_outcomes())

# Prepare the recipe
lasso_prep <- lasso_rec %>%
  prep()
```
Now I will impliment the function. This involves setting the specific LASSO regression engine, creating a workflow, and then extracting the fit of the model.
```{r}
# Specify logistic regression model
lasso_spec <- logistic_reg(penalty = 0.05, mixture = 1) %>%
  set_engine("glmnet")

# Create workflow
lasso_wf <- workflow() %>%
  add_recipe(lasso_rec)

# Fit the model using training data
lasso_fit <- lasso_wf %>%
  add_model(lasso_spec) %>%
  fit(data = train_data)

# Extract and tidy model parameters
lasso_fit %>%
  extract_fit_parsnip() %>%
  tidy()
```

Hi Kelly, I am confused because only the intercept has an estimate, while every other factor is zero. I wonder if maybe the penalty is too high? I am not sure how to fix this.

# Tuning LASSO model