### 2024-04-05 code/analysis-code/analysis3.qmd
Finished analysis 3

---
title: "Data Analysis and Modeling of Teen Vaccination Surveys"
date: '`r format(Sys.Date())`'
Author: Kelly Cao and Rachel Robertson
output: html_document
editor: 
  markdown: 
    wrap: sentence
---
# LASSO using Tidymodels
```{r setup, include=FALSE} 
knitr::opts_chunk$set(warning = FALSE, message = FALSE) 
```
We will start by opening the libraries that we will need for creating a LASSO regression model for the data
```{r}
library(tidymodels) # use tidymodels framework
library(ggplot2) # producing visual displays of data
library(dplyr) # manipulating and cleaning data
library(here) # making relative pathways
library(glmnet) # for LASSO regression engine
library(doParallel) # for parallel processing
library(rsample) # for cross validation
library(yardstick) # for metrics
library(mlbench)

```

Now, we will load the preprocessed data.
```{r}
# Load and preprocess data
data_location <- here::here("data","processed-data","cleandata1.rds")
mydata <- readRDS(data_location)

# Remove rows with missing values
mydata$SEX <- droplevels(mydata$SEX, exclude = c("DON'T KNOW", "MISSING IN ERROR", "REFUSED"))
mydata$INS_STAT2_I <- droplevels(mydata$INS_STAT2_I, exclude = "MISSING Data")
mydata$STATE <- droplevels(mydata$STATE, exclude = "Missing Data")
mydata$MOBIL_1 <- droplevels(mydata$MOBIL_1, exclude = c("DON'T KNOW", "MISSING IN ERROR", "REFUSED"))
mydata$FACILTY <- droplevels(mydata$FACILITY, exclude = "Missing Data")
mydata$P_UTDHPV <- droplevels(mydata$P_UTDHPV, exclude = "Missing Data")
```
Next, we will split the data into training and testing groups so that the LASSO regression has a comparison group.
```{r}
# Split data into training and testing datasets
set.seed(123) # seed for reproducibility
split_data <- initial_split(mydata, prop = 0.8) # 80% split for training/testing data
train_data <- training(split_data)
test_data <- testing(split_data)
```
Next, we will specify the LASSO recipe, or the regression formula. This will contain the predictors of interest used for previous analysis and `P_UTDHPV` as the outcome. 
```{r}
# Create recipe containing predictors of interest
lasso_rec <- recipe( P_UTDHPV ~ AGE + SEX + STATE + INS_STAT2_I + INCQ298A + INS_BREAK_I + INCPOV1 + RACEETHK + EDUC1 + LANGUAGE + MOBIL_1 + RENT_OWN + FACILITY, data = train_data) %>%
  step_dummy(all_nominal(), -all_outcomes())  %>% # convert categorical variables (nominal) into dummy variables
  step_rm(all_outcomes(), -all_outcomes()) #remove the outcome variables from the dataset

# Prepare the recipe
lasso_prep <- lasso_rec %>%
  prep()
```
Now I will implement the function. This involves setting the specific LASSO regression engine, creating a workflow, and then extracting the fit of the model.
```{r}
# Specify logistic regression model
lasso_spec <- logistic_reg(penalty = 0.05, mixture = 1) %>%
  set_engine("glmnet")

# Create workflow
lasso_wf <- workflow() %>%
  add_recipe(lasso_rec)

# Fit the model using training data
lasso_fit <- lasso_wf %>%
  add_model(lasso_spec) %>%
  fit(data = train_data)

# Extract and tidy model parameters
lasso_fit %>%
  extract_fit_parsnip() %>%
  tidy()

# Extract and tidy model parameters
lasso_coef <- lasso_fit %>%
  extract_fit_parsnip() %>%
  tidy()

# Filter out predictors with non-zero coefficients
lasso_nonzero_coef <- lasso_coef %>%
  filter(estimate != 0)

# View the remaining predictors with non-zero coefficients
num_observations <- nrow(lasso_nonzero_coef)
print(num_observations)

```

It seems like the LASSO model is penalizing the coefficients too heavily, resulting in most of them being shrunk to zero. We are going to try lowering the penalty parameter (penalty) to see if it is effective in solving our problem.

```{r}
# Specify logistic regression model
lasso_spec <- logistic_reg(penalty = 0.005, mixture = 1) %>%
  set_engine("glmnet")

# Create workflow
lasso_wf <- workflow() %>%
  add_recipe(lasso_rec)

# Fit the model using training data
lasso_fit <- lasso_wf %>%
  add_model(lasso_spec) %>%
  fit(data = train_data)

# Extract and tidy model parameters
lasso_fit %>%
  extract_fit_parsnip() %>%
  tidy()

# Extract and tidy model parameters
lasso_coef <- lasso_fit %>%
  extract_fit_parsnip() %>%
  tidy()

# Filter out predictors with non-zero coefficients
lasso_nonzero_coef <- lasso_coef %>%
  filter(estimate != 0)

# View the remaining predictors with non-zero coefficients
num_observations <- nrow(lasso_nonzero_coef)
print(num_observations)

```

```{r}
# Generate predictions
predictions <- lasso_fit %>%
  predict(new_data = mydata) %>%
  bind_cols(mydata)

# Calculate ROC AUC
predictions %>%
  mutate(predicted_numeric = as.numeric(.pred_class)) %>%
  roc_auc(P_UTDHPV, predicted_numeric)

# Generate predictions
predictions2 <- lasso_fit %>%
  predict(new_data = train_data) %>%
  bind_cols(train_data)

# Calculate ROC AUC
predictions2 %>%
  mutate(predicted_numeric = as.numeric(.pred_class)) %>%
  roc_auc(P_UTDHPV, predicted_numeric)

```
When attempting to determine how lowering the penalty effected the ability of the model, we produced some predictions with both the original data set and training data, both performed fairly poorly, with a ROC_AUC score close to 0.5, suggesting the the model created predictions that were comparable to random.

To determine the ideal penalty, we are going to tune the model.

# Tuning LASSO model
Any penalty above 0.05 resulted in all of the predictor coefficient reduced to zero. When performing the following tuning, I set the penalty grid from 0.001 to 0.05. 
```{r}
# Define a grid of penalty values to tune over
penalty_grid <- expand.grid(penalty = seq(0.001, 0.05, length.out = 20))

# Set up parallel processing for faster tuning
registerDoParallel()

# LASSO model specification
lasso_model_tuned <- 
  logistic_reg(penalty = tune(), mixture = 1) %>%
  set_engine("glmnet") %>%
  set_mode("classification")  # Ensure correct mode for classification problem

# LASSO Workflow
lasso_workflow_tuned <- 
  workflow() %>%
  add_model(lasso_model_tuned) %>%
  add_recipe(lasso_rec)

# Set up the grid for tuning
lasso_tuned <- 
  lasso_workflow_tuned %>%
  tune_grid(
    resamples = vfold_cv(train_data, v = 5, repeats = 5), 
    grid = penalty_grid,
    control = control_grid(verbose = TRUE)
  )

# View tuning results
tuned_plot <- autoplot(lasso_tuned)
tuned_plot + ggtitle( "Tuned LASSO Model")

# Collect tuning results
tuning_results <- lasso_tuned %>%
  collect_metrics()

# Filter for ROC AUC metric
roc_auc <- tuning_results %>%
  filter(.metric == "roc_auc")

# View the ROC AUC values
print(roc_auc)

```

When viewing the metrics of the tuning, we can see that the regularization is directly correlated with the reduction in roc_auc and accuracy and the increase in brier_Class. The accuracy is later shown to start increasing suggesting underfitting starts occurring around 0.01. Just reviewing the roc_auc alone, the best performance occurs when the penalty is at its lowest.