### 2024-04-05 code/analysis-code/analysis2.5.qmd
Random decision tree and improving LASSO
RF tree is in analysis 2.5 document and LASSO improved at bottom of analysis3 document. Maybe cross validation tuning and refitting, like we did in class, would make this model the best. RF tree did not turn out well, but I did not use it to make predictions yet.

---
title: "Data Analysis and Modeling of Teen Vaccination Surveys"
date: '`r format(Sys.Date())`'
Author: Kelly Cao and Rachel Robertson
output: html_document
editor: 
  markdown: 
    wrap: sentence
---
# Random Forest Model
```{r}
library(tidymodels) # use tidymodels framework.
library(ggplot2) # producing visual displays of data
library(dplyr) # manipulating and cleaning data
library(here) # making relative pathways
library(randomForest) # making random forest model
library(doParallel) # for parallel processing
library(rsample) # for cross validation
```

```{r}
# Load and preprocess data
data_location <- here::here("data","processed-data","cleandata1.rds")
mydata <- readRDS(data_location)

# Remove rows with missing values
mydata$SEX <- droplevels(mydata$SEX, exclude = c("DON'T KNOW", "MISSING IN ERROR", "REFUSED"))
mydata$INS_STAT2_I <- droplevels(mydata$INS_STAT2_I, exclude = "MISSING Data")
mydata$STATE <- droplevels(mydata$STATE, exclude = "Missing Data")
mydata$MOBIL_1 <- droplevels(mydata$MOBIL_1, exclude = c("DON'T KNOW", "MISSING IN ERROR", "REFUSED"))
mydata$FACILTY <- droplevels(mydata$FACILITY, exclude = "Missing Data")
mydata$P_UTDHPV <- droplevels(mydata$P_UTDHPV, exclude = "Missing Data")
```

```{r}
# Split data into training and testing datasets
set.seed(123) # seed for reproducibility
split_data <- initial_split(mydata, prop = 0.8) # 80% split for training/testing data
train_data <- training(split_data)
test_data <- testing(split_data)
```

```{r}
rf_rec <- recipe(P_UTDHPV ~ AGE + SEX + STATE + INS_STAT2_I + INCQ298A + INS_BREAK_I + INCPOV1 + RACEETHK + EDUC1 + LANGUAGE + MOBIL_1 + RENT_OWN + FACILITY, data = train_data) # use full model as recipe for the random forest model

rf_model <- rand_forest()%>% # use rand_forest() to make a random forest model
  set_engine("randomForest", seed = 123)%>%
  set_mode("classification")

rf_workflow <- workflow() %>% # create workflow for rf model
  add_recipe(rf_rec)%>% # apply recipe
  add_model(rf_model) # apply model

rf_fit <- rf_workflow%>%
  fit(data = test_data)%>% # use the workflow to fit the rf model to the data
print(rf_fit)
```

```{r}

tune_spec <- 
  rand_forest(
    mtry = tune(), # parameters of random forest model to tune are the mtry, trees, and min_n
    trees = 50,
    min_n = tune()) %>% 
  set_engine("randomForest", seed = 123) %>% # make sure to set seed again for the tree
  set_mode("classification")

## I used ChatGPT to help me define the tree grid for the parameters mtry and min_n
tree_grid <- grid_regular(mtry(range = c(1, 4)), 
                          min_n(range = c(1, 5)),
                          levels = 6)

rf_tune_wf <- workflow() %>% # create workflow
  add_recipe(rf_rec) %>% # add recipe from earlier
  add_model(tune_spec) # add spec that is defined above

folds_data <- vfold_cv(test_data, v=5 )
# Set the number of cores to use
num_cores <- detectCores() - 1

# Initialize parallel backend
doParallel::registerDoParallel(cores = num_cores)

set.seed(123) # set seed for reproducibility

rf_res <- rf_tune_wf %>%
  tune_grid(
    resamples = folds_data, # use data from the CV folds
    grid = tree_grid)
rf_res %>%
  autoplot()

# Stop parallel processing
stopImplicitCluster()
```