### 2024-04-13 code/analysis-code/analysis2.5.qmd
Added to RF model
RF model added and tuned.

---
title: "Data Analysis and Modeling of Teen Vaccination Surveys"
date: '`r format(Sys.Date())`'
Author: Kelly Cao and Rachel Robertson
output: html_document
editor: 
  markdown: 
    wrap: sentence
---
# Random Forest Model
```{r}
library(tidymodels) # use tidymodels framework.
library(ggplot2) # producing visual displays of data
library(dplyr) # manipulating and cleaning data
library(here) # making relative pathways
library(randomForest) # making random forest model
library(doParallel) # for parallel processing
library(rsample) # for cross validation
library(yardstick)
```

```{r}
# Load and preprocess data
data_location <- here::here("data","processed-data","cleandata1.rds")
mydata <- readRDS(data_location)

# Remove rows with missing values
mydata$SEX <- droplevels(mydata$SEX, exclude = c("DON'T KNOW", "MISSING IN ERROR", "REFUSED"))
mydata$INS_STAT2_I <- droplevels(mydata$INS_STAT2_I, exclude = "MISSING Data")
mydata$STATE <- droplevels(mydata$STATE, exclude = "Missing Data")
mydata$MOBIL_1 <- droplevels(mydata$MOBIL_1, exclude = c("DON'T KNOW", "MISSING IN ERROR", "REFUSED"))
mydata$FACILTY <- droplevels(mydata$FACILITY, exclude = "Missing Data")
mydata$P_UTDHPV <- droplevels(mydata$P_UTDHPV, exclude = "Missing Data")
```

```{r}
# Split data into training and testing datasets
set.seed(123) # seed for reproducibility
split_data <- initial_split(mydata, prop = 0.8) # 80% split for training/testing data
train_data <- training(split_data)
test_data <- testing(split_data)
```

```{r}
rf_rec <- recipe(P_UTDHPV ~ AGE + SEX + STATE + INS_STAT2_I + INCQ298A + INS_BREAK_I + INCPOV1 + RACEETHK + EDUC1 + LANGUAGE + MOBIL_1 + RENT_OWN + FACILITY, data = train_data) # use full model as recipe for the random forest model

rf_model <- rand_forest()%>% # use rand_forest() to make a random forest model
  set_engine("randomForest", seed = 123)%>%
  set_mode("classification")

rf_workflow <- workflow() %>% # create workflow for rf model
  add_recipe(rf_rec)%>% # apply recipe
  add_model(rf_model) # apply model

rf_fit <- rf_workflow%>%
  fit(data = train_data)%>% # use the workflow to fit the rf model to the data
print(rf_fit)
```
Now, I will examine the performance of the base random forest model.
```{r}
# Performance metrics of the first fit
rf_aug <- augment(rf_fit, train_data) # augment to make predictions for rf_fit

metrics <- metric_set(accuracy, f_meas) # create a set of metrics to test for classification
first_fit_metrics <- metrics(truth = P_UTDHPV, estimate = .pred_class, data = rf_aug) # calculate metrics for first fit
print(first_fit_metrics)
```
The accuracy of the model is fairly high, meaning that there it performs fairly when classifying the predictors in the model. However, the F1 score is relatively low which reflects low precision and recall of the model. To improve this we will tune the model.

Now, I will continue by tuning the RF model using cross validation. I start with a smaller numer of trees and lower number of folds because my computer cannot computationally handle many trees or folds.
```{r}
tune_spec <- 
  rand_forest(
    mtry = tune(), # parameters of random forest model to tune are the mtry, trees, and min_n
    trees = 100,
    min_n = tune()) %>% 
  set_engine("randomForest", seed = 123) %>% # make sure to set seed again for the tree
  set_mode("classification")

## I used ChatGPT to help me define the tree grid for the parameters mtry and min_n
tree_grid <- grid_regular(mtry(range = c(1, 13)), 
                          min_n(range = c(1, 200)),
                          levels = 6)

rf_tune_wf <- workflow() %>% # create workflow
  add_recipe(rf_rec) %>% # add recipe from earlier
  add_model(tune_spec) # add spec that is defined above

folds_data <- vfold_cv(train_data, v=5 )
# Set the number of cores to use
num_cores <- detectCores() - 1

# Initialize parallel backend
doParallel::registerDoParallel(cores = num_cores)

set.seed(123) # set seed for reproducibility

rf_res <- rf_tune_wf %>%
  tune_grid(
    resamples = folds_data, # use data from the CV folds
    grid = tree_grid)
rf_res %>%
  autoplot()

# Stop parallel processing
stopImplicitCluster()
```
```{r}
# Find the best values for the tuned metrics, mtry and min_n
# find the ideal ROC_AUC metric for the rf tree
rf_res %>%
  collect_metrics() %>% # get metrics from the rf model created
  filter(.metric == "roc_auc") %>% # filter for the roc auc metric
  select(mean, min_n, mtry) %>% # select the columsn for mean, min_n and mtry
  pivot_longer(min_n:mtry,
    values_to = "value",
    names_to = "parameter"
  ) %>%
# Plot the ROC_AUC to find the ideal min_n and mtry
  ggplot(aes(value, mean, color = parameter)) +
  geom_point(show.legend = FALSE) +
  facet_wrap(~parameter, scales = "free_x") +
  labs(x = NULL, y = "AUC")
```
It seems that mtry is maximized between 100 and 150 and min_n is maximized between 1 and 5.
We can use a grid search to tune within a range to determine which is the ideal parameter value.

```{r}
# grid search tuning
rf_tune_grid <- grid_regular( #define a grid with a range for mtry and min_n
  mtry(range = c(100, 150)),
  min_n(range = c(1, 5)),
  levels = 6
)

rf_tune_grid # display the grid

# Tune using the defined tuning grid
doParallel::registerDoParallel(cores = num_cores)

set.seed(123) # set seed for reproducibility

rf_res2 <- rf_tune_wf %>%
  tune_grid(
    resamples = folds_data, # use data from the CV folds
    grid = rf_tune_grid) # use new tune grid
rf_res2 %>%
  autoplot()

# Stop parallel processing
stopImplicitCluster()
```
From the plot, it seems that the ideal minimum node size (min_n) is 5 when the number of randomly selected predictors (mtry) is  140, when accuracy and ROC_AUC are both accounted for. We can check this using the select_best function.
```{r}
# Specify the metric and optimization criteria
best_auc <- select_best(rf_res2, metric = "roc_auc")

final_rf <- finalize_model( # select the best model based on roc_auc
  tune_spec,
  best_auc
)
# fit the tuned model
rf_rec <- recipe(P_UTDHPV ~ AGE + SEX + STATE + INS_STAT2_I + INCQ298A + INS_BREAK_I + INCPOV1 + RACEETHK + EDUC1 + LANGUAGE + MOBIL_1 + RENT_OWN + FACILITY, data = train_data) # use full model as recipe for the random forest model

rf_model <- rand_forest()%>% # use rand_forest() to make a random forest model
  set_engine("randomForest", seed = 123)%>%
  set_mode("classification")

rf_workflow <- workflow() %>% # create workflow for rf model
  add_recipe(rf_rec)%>% # apply recipe
  add_model(rf_model) # apply model

rf_fit <- rf_workflow%>%
  fit(data = train_data)%>% # use the workflow to fit the rf model to the data
print(rf_fit)
```
The final random forest model was selected to optimize roc_auc, but it also has a fairly high accuracy. We can use
```{r}
final_rf_fit <- fit(final_rf, data = train_data)

```

```{r}
# RF model compared to single predictor model
## Create a single model
rf_rec_null <- recipe(P_UTDHPV ~ SEX, data = train_data) # specify the model with a recipe; I could not get a null model to work so I used one predictor as a minimal model for comparison

rf_spec_null <- rand_forest() %>%
  set_engine("randomForest", importance = TRUE, seed = 123) %>% # set the seed and specify that importance is true
  set_mode("classification")

rf_workflow_mull <- workflow() %>% # create workflow with the recipe and spec
  add_recipe(rf_rec_null) %>% 
  add_model(rf_spec_null)

rf_fit_null <- rf_workflow_null %>% # fit the null model
  fit(data = train_data)

# Get predictions for tuned model
rf_tuned <- augment(rf_res, train_data)

# Get predictions for first fit model

# Create combined data frame

# Plot tuned model, first fit, and single predictor model together

```
```{r}
# RF model performance with cross validation

```