### 2024-04-14 code/analysis-code/analysis4.qmd
Updated analysis

---
title: "Data Analysis and Modeling of Teen Vaccination Surveys"
date: '`r format(Sys.Date())`'
Author: Kelly Cao and Rachel Robertson
output: html_document
editor: 
  markdown: 
    wrap: sentence
---
# Finding Best Model
Comparing models