### 2024-04-17 code/analysis-code/analysis4.qmd

---
title: "Data Analysis and Modeling of Teen Vaccination Surveys"
date: '`r format(Sys.Date())`'
Author: Kelly Cao and Rachel Robertson
output: html_document
editor: 
  markdown: 
    wrap: sentence
---
# Finding Best Model
Comparing the previous three models used in analysis1.qmd, analysis2.5.qmd, and analysis4.qmd. These models include : general logistic regression model, Random Forest regression model, and LASSO regression model.

# Summary of the predictions and Metrics

# Test Data