### 2024-04-29 code/analysis-code/analysis4.qmd
Made elastic net model and tested
tuned elastic net model performs the same as the previous lasso model. Because of this, I am thinking that we should choose the rf model with the best performance metrics (despite being overfit).

---
title: "Data Analysis and Modeling of Teen Vaccination Surveys"
date: '`r format(Sys.Date())`'
Author: Kelly Cao and Rachel Robertson
output: html_document
editor: 
  markdown: 
    wrap: sentence
---
# Elastic Net Regression
Because the random forest model was overfit and the lasso model did not capture all of the trends, we will create an elastic net model of our categorical variables.

## Loading data
We will begin by loading the data and opening the libraries needed for the analysis
```{r}
# Open the libraries needed
library(tidymodels) # use tidymodels framework
library(ggplot2) # producing visual displays of data
library(dplyr) # manipulating and cleaning data
library(here) # making relative pathways
library(glmnet) # for LASSO regression engine
library(doParallel) # for parallel processing
library(rsample) # for cross validation
library(yardstick) # for metrics
library(mlbench)
library(vip)

# Load and preprocess data
data_location <- here::here("data","processed-data","cleandata1.rds")
mydata <- readRDS(data_location)

# Remove rows with missing values
mydata$SEX <- droplevels(mydata$SEX, exclude = c("DON'T KNOW", "MISSING IN ERROR", "REFUSED"))
mydata$INS_STAT2_I <- droplevels(mydata$INS_STAT2_I, exclude = "MISSING Data")
mydata$STATE <- droplevels(mydata$STATE, exclude = "Missing Data")
mydata$MOBIL_1 <- droplevels(mydata$MOBIL_1, exclude = c("DON'T KNOW", "MISSING IN ERROR", "REFUSED"))
mydata$FACILTY <- droplevels(mydata$FACILITY, exclude = "Missing Data")
mydata$P_UTDHPV <- droplevels(mydata$P_UTDHPV, exclude = "Missing Data")
```
Next, we will split the data into training and testing groups so that the LASSO regression has a comparison group.
```{r}
# Split data into training and testing datasets
set.seed(123) # seed for reproducibility
split_data <- initial_split(mydata, prop = 0.8) # 80% split for training/testing data
train_data <- training(split_data)
test_data <- testing(split_data)
```
## Elastic net first fit
Now we will proceed by fitting the elastic ent model for the first time
```{r}
# Split the data into training and testing sets
set.seed(123)  # for reproducibility

# Create amodel with the specifications for elatic net
elastic_net <- logistic_reg(mode = "classification") %>%
  set_engine("glmnet") %>%
  set_mode("classification") %>%
  set_args(
    penalty = 0.005,  # use the penalty for the previous lasso regression
    mixture = 0.5   
  )

# create a recipe, using your formula, for elastic net
elastic_rec <- recipe(P_UTDHPV ~ AGE + SEX + STATE + INS_STAT2_I + INCQ298A + INS_BREAK_I + INCPOV1 + RACEETHK + EDUC1 + LANGUAGE + MOBIL_1 + RENT_OWN + FACILITY, data = train_data) %>%
  step_dummy(all_nominal(), -all_outcomes())  %>% # convert categorical variables (nominal) into dummy variables
  step_rm(all_outcomes(), -all_outcomes()) #remove the outcome variables from the dataset

# Prepare the elastic net recipe
elastic_prep <- elastic_rec %>%
  prep()

# produce elastic net workflow
workflow <- workflow() %>%
  add_recipe(elastic_rec) %>%
  add_model(elastic_net)

# implement elastic net workflow
en_fit <- workflow %>% 
  fit(data = train_data) # fit to training data

en_predictions <- augment(en_fit, train_data) #make predictions on training data
print(en_predictions)
```
Now that we have produced predictions with the elastic net model, we will look at the performance metrics of this model and then tune the model parameters.
```{r}
# Calculate ROC AUC
roc_auc <- en_predictions %>%
  roc_auc(truth = P_UTDHPV, .pred_UTD)
print(roc_auc)
```
The ROC AUC reveals that there is no predictive power with the elastic net model, so I will tune this model to reveal if it produces better predictions.
```{r}
# Tune for penalty and mixture parameters

# produce the model spec using the same recipe, except with the aprameters for tuning using tune()
elastic_spec <- logistic_reg(mode = "classification") %>%
  set_engine("glmnet") %>%
  set_mode("classification") %>%
  set_args(
    penalty = tune(), #set both penalty and mixture parameters to tune  
    mixture = tune()
  )
# make the tuning grid with the range for penalty and for mixture values
tune_grid <- expand.grid(
  penalty = seq(0.001, 0.05, length.out = 20),
  mixture = c(0.25, 0.5, 0.75)
)

folds <- vfold_cv(train_data, v = 5, repeats = 5) # define number of folds and repeats for cross validation

# Set up parallel processing for faster tuning
registerDoParallel()

set.seed(123) # set seed for predictability

# produce elastic net workflow using the tuning spec
workflow <- workflow() %>%
  add_recipe(elastic_rec) %>%
  add_model(elastic_spec) #use the tuned spec for the model

en_tuned <- 
  workflow %>%
  tune_grid(
    resamples = folds, 
    grid = tune_grid,
    control = control_grid(verbose = TRUE)
  )

# View tuning results
tuned_plot <- autoplot(en_tuned)
tuned_plot + ggtitle( "Tuned elastic Net Model")

# Collect tuning results
tuning_results <- en_tuned %>%
  collect_metrics()

# Filter for ROC AUC metric
roc_auc <- tuning_results %>%
  filter(.metric == "roc_auc")

# View the ROC AUC values
print(roc_auc)
```
For the results of the three graphs, we can see that the ideal mixture is likely 0.25. The amount of regularization, is best around 0.01, for a mixture of 0.25. Because of this, I will select the best model with a mixture of 0.25 and penalty of ~ 0.01.

```{r}
# select the model based on the best roc auc score
best_auc <- select_best(en_tuned, metric = "roc_auc")

final_en <- finalize_model( # finalize the model based on the roc auc score
  elastic_spec,
  best_auc
)
# Split the data into training and testing sets
set.seed(123)  # for reproducibility

# produce elastic net workflow
workflow <- workflow() %>%
  add_recipe(elastic_rec) %>%
  add_model(final_en)

# implement elastic net workflow
en_tune_fit <- workflow %>% 
  fit(data = train_data) # fit to training data

en_predictions_tune <- augment(en_tune_fit, train_data) #make predictions on training data
print(en_predictions)

# Calculate ROC AUC
roc_auc <- en_predictions_tune %>%
  roc_auc(truth = P_UTDHPV, .pred_UTD)
print(roc_auc)
```
The ROC AUC has slightly improved, to 0.335 rather then 0.338. I will now evaluate other performance metrics of the model, including rmse and accuracy to see if the model has improved.
```{r}
tuned_metrics <- metrics(truth = P_UTDHPV, estimate = .pred_class, data = en_predictions) # calculate metrics for first fit
print(tuned_metrics)
```
The accuracy of the model is fairly high, at 0.79 or 79%, but we will see how these performance metrics change after using the test data to make predictions.
```{r}
# use test data ot make predictions
en_predictions_test <- augment(en_tune_fit, test_data) #make predictions on testing data
print(en_predictions_test)

# evaluate roc_auc
roc_auc <- en_predictions_test %>%
  roc_auc(truth = P_UTDHPV, .pred_UTD)
print(roc_auc)

#evaluate accuracy and kap
tuned_metrics <- metrics(truth = P_UTDHPV, estimate = .pred_class, data = en_predictions_test) # calculate metrics for first fit
print(tuned_metrics)

# evaluate the F1 score of the model predictions
f1 <- en_predictions_test %>%
  f_meas(truth = P_UTDHPV, estimate = .pred_class)
print(f1)
```
The ROC_AUC stayed fairly similar, at 0.365 and the accuracy remained the same, around 0.79 and a kap value of around 0.036. The kappa value is best around 1 or -1, so a kap of 0.036 is not good.

Despite this, I will see if the important predictors are in agreement with our LASSO model using the vip package.
```{r}
# Finding importance of variables chosen
en_vars <- en_tune_fit %>%
  extract_fit_parsnip() %>% # use parsnip to extract to fit
  vip::vip() #use vip for visualizing the most important variables

print(en_vars)
```
The elastic net model produced the same important variables as the previous lasso model.