### 2024-05-03 code/analysis-code/analysis2.qmd

---
title: "Data Analysis and Modelling of Teen Vaccination Surveys"
date: '`r format(Sys.Date())`'
Author: Kelly Cao and Rachel Robertson
output: html_document
editor: 
  markdown: 
    wrap: sentence

---
### Introduction
The previous analysis performed shown on analysis1.qmd found in the same directory, showed collinearity among many of the predictor variables. The following script is an attempt to reduce that issue to prevent problem when performing analyses further downstream.

```{r setup, include=FALSE} 
knitr::opts_chunk$set(warning = FALSE, message = FALSE) 
```

#### Loading library and data set
``` {R, warnings=FALSE, messages=FALSE}
# Load required libraries
library(MASS) # for robust regression
library(caret) # for cross-validation
library(h2o) #For dimension reduction
library(dplyr) 
library(rsample)
library(knitr)
library(kableExtra)
library(webshot)
library(here)
library(gt)
library(gtsummary)
```

I start with loading the same data set and performing the same functions to clean up the data by removing NAs. We will be looking at the variable P_UTDHPV, which reflects the up-to-date HPV vaccination status of the individual.
```{R}
# Load and preprocess data
data_location <- here::here("data","processed-data","cleandata1.rds")
mydata <- readRDS(data_location)

# Remove rows with missing values for the response variable
mydata <- mydata[complete.cases(mydata$P_UTDHPV), ]

# Set Seed for reproducibility
rngseed = 1234
set.seed(rngseed)

# Create training data by alloting 3/4 of the data for training
data_split <- initial_split(mydata, prop = 3/4)

# Create data frames for the two sets:
train_data <- training(data_split)
test_data  <- testing(data_split)

```

### PCA Analysis
I then initialize the h2o cluster and converted my data object to a H2oFrame or an h2o object. I then performed PCA using the function `h20.prcomp` and using the newly converted h20 object as a training frame.

This is perform in attempt t oindirectly reduce the collinearity in the predictor variables.

The following lines are based on the resource shown below:

[Link](https://bradleyboehmke.github.io/HOML/pca.html)

>PCA indirectly addresses multicollinearity by transforming the original features into a new set of orthogonal variables, whereas LASSO/Ridge regression directly addresses multicollinearity by penalizing the coefficients of correlated predictors.

```{R}
# Initialize and connect to the H2O cluster
# Start an H2O cluster
h2o.init()

# Convert train_data to H2OFrame
mydata_h2o <- as.h2o(train_data)

# Perform PCA using h2o.prcomp
pca_model <- h2o.prcomp(
  training_frame = mydata_h2o, 
  x = c("AGE", "SEX", "STATE", 
        "INS_STAT2_I", "INCQ298A", 
        "INS_BREAK_I", "INCPOV1", 
        "RACEETHK", "EDUC1", 
        "LANGUAGE", "MOBIL_1", 
        "RENT_OWN", "FACILITY"))

# Viewing the result of the PCA
print(pca_model)

```

```{r}
# Extract the components from the PCA model
standard_deviation <- pca_model@model$model_summary$pc1[1]
proportion_variance <- pca_model@model$model_summary$pc1[2]
cumulative_proportion <- pca_model@model$model_summary$pc1[3]

# Create a data frame with the PC information
scree_data <- data.frame(
  Component = "PC1",
  Standard.deviation = standard_deviation,
  Variance = standard_deviation^2,
  Proportion.Variance = proportion_variance,
  Cumulative.Proportion = cumulative_proportion
)

# Print the summary table
print(scree_data)

# Create a table using kableExtra
pca_table <- 
  scree_data %>%
  gt() %>%
  tab_header(
    title = "PCA Summary Table"
  )

# Save the table to a file
gtsave(pca_table, here("results","tables", "pca_summary.png"))
```

The PCA results showed one principle component with a high standard deviation and high proportion of variance, showing that the Pc1 captures a large portion (97.9%) of variability in the data.

The result of having only one principle component further supports the fact that many of the predictor variables are highly collinear. One PC shows that a PCA must reduce the data object to a single dimension. Having a single principal component could be problematic for my analysis, particularly if it leads to the loss of crucial information or structure in my data due to collinearity among the original features.

Understanding why this occurs and assessing whether it aligns with my expectations and the characteristics of my dataset is essential. I may need to potentially review my data preprocessing methods or explore different feature selection techniques.

### LASSO Regression
I first investigate alternative approaches to address collinearity, such as employing regularization methods like LASSO regression. This would ensure that I'm effectively capturing the underlying patterns in my data while mitigating the challenges posed by collinearity.

```{R}
# Use LASSO regression for variable selection and regularization
lasso_model <- h2o.glm(
  x = c("AGE", "SEX", "STATE",
        "INS_STAT2_I", "INCQ298A"
        , "INS_BREAK_I", "INCPOV1",
        "RACEETHK", "EDUC1",
        "LANGUAGE", "MOBIL_1",
        "RENT_OWN", "FACILITY"),
  y = "P_UTDHPV", training_frame = mydata_h2o, family = "binomial", alpha = 1)

# View Lasso Model
print(lasso_model)

```

The model used Lasso to pick out 97 important predictors from 105, which likely dealt with issues of collinearity among the original features. This simplification helped in making the model easier to understand and potentially improved its performance.

### Attempt at Cross-validation
I then utilized ChatGPT to write me a code to perform a cross-validation test.

```{R}
# Create a fold column for cross-validation
fold_column <- h2o.kfold_column(data = mydata_h2o, nfolds = 10, seed = 1)  # Specify the number of folds and a seed for reproducibility

# Add the fold column to the H2OFrame
mydata_h2o <- h2o.cbind(mydata_h2o, fold_column)

# Specify predictor variables
predictor_columns <- c("AGE", "SEX", "STATE", "INS_STAT2_I", "INCQ298A", "INS_BREAK_I", 
                       "INCPOV1", "RACEETHK", "EDUC1", "LANGUAGE", "MOBIL_1", "RENT_OWN", "FACILITY")

# Evaluate model performance using cross-validation
cv_metrics <- h2o.glm(x = predictor_columns, 
                      y = "P_UTDHPV", 
                      training_frame = mydata_h2o, 
                      fold_column = "C1",  # Use the name of the fold column (default name is "C1")
                      family = "binomial")

# Display results
print(cv_metrics)
```

The cross-validation results for the logistic regression model show consistent performance across folds, with an average accuracy of approximately 79.4%. The model demonstrates stability with metrics such as an area under the curve (AUC) around 0.644 and a mean per-class error of about 49.8%. Precision and recall metrics also exhibit robustness, averaging around 79.6% and 99.6%, respectively. These results indicate the model's ability to balance correctly identifying positive cases while minimizing false positives. However, areas for model refinement include residual deviance and specificity.

```{r}
# Replicating the previous analysis with test data

# Convert train_data to H2OFrame
mydata_h2o <- as.h2o(test_data)

# Create a fold column for cross-validation
fold_column <- h2o.kfold_column(data = mydata_h2o, nfolds = 10, seed = 1)  # Specify the number of folds and a seed for reproducibility

# Add the fold column to the H2OFrame
mydata_h2o <- h2o.cbind(mydata_h2o, fold_column)

# Specify predictor variables
predictor_columns <- c("AGE", "SEX", "STATE", "INS_STAT2_I", "INCQ298A", "INS_BREAK_I", 
                       "INCPOV1", "RACEETHK", "EDUC1", "LANGUAGE", "MOBIL_1", "RENT_OWN", "FACILITY")

# Evaluate model performance using cross-validation
cv_metrics <- h2o.glm(x = predictor_columns, 
                      y = "P_UTDHPV", 
                      training_frame = mydata_h2o, 
                      fold_column = "C1",  # Use the name of the fold column (default name is "C1")
                      family = "binomial")

# Display results
print(cv_metrics)

# Function to extract metrics from a model
extract_metrics <- function(model, dataset) {
  metrics <- model@model$training_metrics@metrics
  mse <- metrics$MSE
  r2 <- metrics$r2
  logloss <- metrics$logloss
  mpce <- metrics$mean_per_class_error
  auc <- metrics$AUC
  gini <- metrics$Gini
  nulldev <- metrics$null_deviance
  resdev <- metrics$residual_deviance
  aic <- metrics$AIC

  # Create a data frame for metrics
  metrics_df <- data.frame(
    Metric = c("MSE", "R^2", "Log Loss", "Mean Per-Class Error", "AUC", "Gini", "Null Deviance", "Residual Deviance", "AIC"),
    Value = c(mse, r2, logloss, mpce, auc, gini, nulldev, resdev, aic),
    Dataset = dataset
  )

  return(metrics_df)
}

# Extract metrics from the LASSO model (training data)
train_metrics_df <- extract_metrics(lasso_model, "Training")

# Extract metrics from the CV metrics (test data)
test_metrics_df <- extract_metrics(cv_metrics, "Test")

# Combine the training and test metrics data frames
combined_metrics_df <- rbind(train_metrics_df, test_metrics_df)

# Create a gt table
gt_table <- combined_metrics_df %>%
  gt(
    groupname_col = "Dataset", 
    row_group_as_column = TRUE
  ) %>%
  tab_header(
    title = md("Model Evaluation Metrics"),
    subtitle = md("Comparison of Metrics between Test and Training Datasets")
  )

# Save the gt table
gtsave(gt_table, here("results", "tables", "LASSO_metrics.png"))

```

### Conclusion
our analysis aimed to address collinearity issues in our teen vaccination surveys dataset through dimension reduction techniques like PCA and regularization methods such as LASSO regression. LASSO regression effectively selected important predictors while simplifying the model. Subsequent cross-validation demonstrated consistent model performance, with an average accuracy of 79.4% and robust precision and recall metrics.