### 2024-05-03 products/manuscript/supplement/Supplementary-Material.qmd

---
title: "Supplementary Material"

author: rachel Robertson and Kelly Cao
date: "`r Sys.Date()`"
format:
  docx:
    toc: false
    number-sections: true
    highlight-style: github
---
```{r, echo=FALSE, message=FALSE}
# load packages
library(here)
library(dplyr)
library(knitr)
library(kableExtra)
```
The supplementary materials include tables and figures from the exploratory analysis and data analysis that were unused in the final manuscript.

![S.1. Distribution of ethnicity and race in sample populaiton of U.S. teens.](../../../results/figures/ethnicity.distribution.png)

![S.2. The distribution of facilities in which the survey was given to study sample of U.S. teens.](../../../results/figures/facility.distribution.png)
![S.3. Distribution of income poverty ratio compared to HPV vaccination status (up-to-date or not up-to-date).](../../../results/figures/income-vaccination.png)

![S.4. distribution of insurance status in study population of U.S. teenagers.](../../../results/figures/insurance.status.distribution.png)
![S.5. The distribution of race/ ethnicity in U.S. teens in the study population, stratified by HPV vaccination status.](../../../results/figures/race-ethnicity-vaccination.png)
![S.6. The distribution of race/ ethnicity in U.S. teens in the study population.](../../../results/figures/race.distribution.png)

![S.7. The percentage of surveys submitted for each state in the U.S. study.](../../../results/figures/state_distribution.png)

![S.8. The distribution of HPV vaccination status in the overall study population of U.S. teens.](../../../results/figures/vaccination.status.distribution.png)

![S.9. The distribution of race/ ethnicity in U.S. teens in the study population, stratified by HPV vaccination status.](../../../results/figures/race-ethnicity-vaccination.png)

```{r, warning= FALSE, message = FALSE, echo = FALSE}
#read the table
table1 <- here("results/tables/state-vaccine-completion.txt")
S10 <- read.table(table1, header = TRUE)

# Render the table using kable
kable(S10, caption = "S.10. Percentage of up-to-date HPV vaccination status for each state from study sample.")
```

```{r, warning= FALSE, message = FALSE, echo = FALSE}
#read the table
table2 <- here("results/tables/vaccine_percent_by_race.txt")
S11 <- read.table(table2, header = TRUE)

# Render the table using kable
kable(S11, caption = "S.11. Percentage and Count of up-to-date HPV vaccination status according to racial and ethnic group.")
```